---
title: "MMP Data Analysis and Models"
author: "Jan Sodoge"
date: "3/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stargazer)
library(broom)
library(DescTools)
```


# Create Sample
Following the criteria outlined in the thesis the dataset is created based on the full 
MMP dataset. 
```{r,  echo=FALSE}
#comun data
commun <- read.csv("commun174.csv")
house <- read.csv("house170.csv")
life <- read.csv("life170.csv")

pers <- read.csv("pers170.csv")
mig <- read.csv("mig170.csv")


commun_filtered <- commun %>% 
  filter(COMPOP90 < 2500) %>% #less than 5000 inhabitants in community --> rural
  filter(AGRIM90 > .5) # more than 50% of men work in agriculture --> agrarian







house_filtered <- house %>% 
  dplyr::semi_join(commun_filtered, by = c("commun" = "COMMUN")) %>%  #investigate only HH in rural agrarian communities
  dplyr::filter(famw != 8888 & famw != 9999 & famw != 0) %>% 
  dplyr::filter(clhect1 != 8888 & clhect1 != 9999) %>% #use only those who own land
  dplyr::filter(clhect2 == 8888 | clhect2 == 9999 | clhect2 < 10) %>% 
  dplyr::filter(clhect3 == 8888 | clhect3 == 9999 | clhect3 < 10) %>% 
  dplyr::filter(clhect4 == 8888 | clhect4 == 9999 | clhect4 < 10) %>% 
  dplyr::filter(clhect1 < 10) %>%   #aim for smallholder agriculture
  mutate(clhect2 = ifelse(clhect2 == 8888 | clhect2 == 9999, 0, clhect2)) %>% 
  mutate(clhect3 = ifelse(clhect3 == 8888 | clhect3 == 9999, 0, clhect3)) %>% 
  mutate(clhect4 = ifelse(clhect4 == 8888 | clhect4 == 9999, 0, clhect4)) %>% 
  mutate(total_land = clhect1 + clhect2 + clhect3, clhect4)







life_filtered <- life %>% 
  semi_join(house_filtered,  by = c("commun" = "commun", "hhnum" = "hhnum", "surveyyr" = "surveyyr")) 
  



pers_filtered <- pers %>% 
  semi_join(house_filtered, by = c("commun" = "commun", "hhnum" = "hhnum", "surveyyr" = "surveyyr"))


mig_filtered <- mig %>% 
  semi_join(house_filtered, by = c("commun" = "commun", "hhnum" = "hhnum", "surveyyr" = "surveyyr"))



```



















# Models predciting migration US
```{r}



life_filtered_us_migrants <- life_filtered %>% 
  dplyr::filter(usyr1 != 8888) %>% 
  filter(usyr1 == year) %>% 
  select(usyr1, fausmig, mousmig, brousmig, sisusmig, land, hectars, age, commun, children) %>% 
  filter(age > 18)


life_filtered_non_migrants <- life_filtered %>% 
  dplyr::filter(usyr1 == 8888) %>% 
  select(usyr1, fausmig, mousmig, brousmig, sisusmig, land, hectars, age, commun, children) %>% 
  filter(age > 18)


age_data_migrants <- as.vector(life_filtered_us_migrants$age)

counterfactual_ages <- data.frame()
for(migrant_age in age_data_migrants){
  tmp <- dplyr::filter(life_filtered_non_migrants, age == migrant_age) 
  
  
  counterfactual_ages <- dplyr::bind_rows(counterfactual_ages, tmp)
  
  
}

counterfactual_ages <- counterfactual_ages %>% 
  distinct()






dataset <- dplyr::bind_rows(life_filtered_us_migrants, counterfactual_ages) %>% 
  mutate(migrated_us = ifelse(usyr1 == 8888, 0, 1),
         social_network = fausmig + mousmig, brousmig + sisusmig,
         children = ifelse(children > 0, 1,0 )) 
  




m1_a <- glm(migrated_us ~ social_network + land, data = dataset, family = "binomial")
m1_b <- glm(migrated_us ~ social_network + hectars, data = dataset, family = "binomial")


m2_a <- glm(migrated_us ~ social_network + land + log(age) + children , data = dataset, family = "binomial")
m2_b <- glm(migrated_us ~ social_network + hectars + log(age) + children , data = dataset, family = "binomial")



m3_a <- glm(migrated_us ~ social_network + land + log(age)+ children +  as.factor(commun), data = dataset, family = "binomial")
m3_b <- glm(migrated_us ~ social_network + hectars + log(age)+ children +  as.factor(commun), data = dataset, family = "binomial")


library(bife)
m3_a_bife <-  bife(migrated_us ~ social_network + land + children + log(age)  | commun, dataset, "probit")
m3_b_bife <-  summary(bife(migrated_us ~ social_network + hectars + children + log(age)  | commun, dataset, "probit"))



stargazer(m1_a, m1_b, m2_a, m2_b, m3_a, m3_b, type = "text")


#prediction_data <- augment(m1, type.predict = "response") %>% 
#  mutate(.pred = as.numeric(.fitted > .5))



#mean(prediction_data$migrated_us == prediction_data$.pred, na.rm = TRUE)


```


```{r}
cat(paste("Derived coefficients used for NetLogo agent-behavior paramtrizstion",
            "
          For intercept", coef(m2_b)[['(Intercept)']],
            "
          For Children", coef(m2_b)[["children"]],
            "
          For social network", coef(m2_b)[["social_network"]],
            "
          For land", coef(m2_b)[["hectars"]],
            "
          For log(age)", coef(m2_b)[["log(age)"]]))
```


Within the Netlogo model this translates to the following structural form to derive a migration probability (which needs to be re-adjusted for out-migration statistics). From the logistic regression above, we derive the following coefficients: 

$\beta_1$ age;
$\beta_2$ children;
$\beta_3$ hectares;
$\beta_4$ network-effect;
$\beta_5$ intercept; 


$$p = exp(\beta_5 + (\beta_1 * log(age)) + (\beta_3 * land) + (\beta_4 * network)) / (1+ exp(\beta_5 + (\beta_1 * log(age)) + (\beta_3 * land) + (\beta_4 * network)))$$






Derive a prediction formula for the above-specified parameters
```{r}
predicted_probs <- expand.grid(age= 20, hectars= seq(0,10), social_network = seq(0,2), children = 1)
estimates <- augment(m2_b, type.predict = "response", newdata = predicted_probs)
estimates <- estimates %>% 
  mutate(social_network = as.factor(social_network))
ggplot(data = estimates, aes(x = hectars, y = .fitted, fill = social_network, color = social_network))+
  geom_line()+
  labs(title="U.S. Migration: Predicted Probabilities")+
  theme_bw()
```






# Models predciting migration Domestic
```{r}
life_filtered_us_migrants <- life_filtered %>% 
  dplyr::filter(doyr1 != 8888) %>% 
  filter(doyr1 == year) %>% 
  select(doyr1, fausmig, mousmig, brousmig, sisusmig, land, hectars, age, commun, children) %>% 
  filter(age > 18)


life_filtered_non_migrants <- life_filtered %>% 
  dplyr::filter(doyr1 == 8888) %>% 
  select(doyr1, fausmig, mousmig, brousmig, sisusmig, land, hectars, age, commun, children) %>% 
  filter(age > 18)


age_data_migrants <- as.vector(life_filtered_us_migrants$age)

counterfactual_ages <- data.frame()
for(migrant_age in age_data_migrants){
  tmp <- dplyr::filter(life_filtered_non_migrants, age == migrant_age) 
  counterfactual_ages <- dplyr::bind_rows(counterfactual_ages, tmp)
}

counterfactual_ages <- counterfactual_ages %>% 
  distinct()



dataset <- dplyr::bind_rows(life_filtered_us_migrants, counterfactual_ages) %>% 
  mutate(migrated_do = ifelse(doyr1 == 8888, 0, 1),
         social_network = fausmig + mousmig, brousmig + sisusmig,
         children = ifelse(children > 0, 1,0 )) 
  




#m1_do <- glm(migrated_do ~  land, data = dataset, family = "binomial")
#m2_do <- glm(migrated_do ~  land + log(age) + children , data = dataset, family = "binomial")
#m3_do <- glm(migrated_do ~ land + log(age)+ children +  as.factor(commun), data = dataset, family = "binomial")
#stargazer(m1_do, m2_do, m3_do, type = "text")


m1_a_do <- glm(migrated_do ~ social_network + land, data = dataset, family = "binomial")
m1_b_do <- glm(migrated_do ~ social_network + hectars, data = dataset, family = "binomial")
m2_a_do <- glm(migrated_do ~ social_network + land + log(age) + children , data = dataset, family = "binomial")
m2_b_do <- glm(migrated_do ~ social_network + hectars + log(age) + children , data = dataset, family = "binomial")
m3_a_do <- glm(migrated_do ~ social_network + land + log(age)+ children +  as.factor(commun), data = dataset, family = "binomial")
m3_b_do <- glm(migrated_do ~ social_network + hectars + log(age)+ children +  as.factor(commun), data = dataset, family = "binomial")


stargazer(m1_a_do, m1_b_do, m2_a_do, m2_b_do, m3_a_do, m3_b_do, type = "text")


```


```{r, results ='asis'}
stargazer(m1_a, m1_b, m2_a, m2_b, m3_a, m3_b,
          m1_a_do, m1_b_do, m2_a_do, m2_b_do, m3_a_do, m3_b_do,
          type = "html")

```









# Receiving remittances module


How many households, where a member migrated, receive remittances? Determines the probability that a household receives remittances


```{r}

us_migrants_currently <- pers %>% 
  filter(uscurtrp == 1) 


remittances_reciving_ratio <- house_filtered %>% 
  semi_join(us_migrants_currently, by = c("commun" = "commun", "hhnum" = "hhnum", "surveyyr" = "surveyyr")) %>%  
  dplyr::filter(remus != 3) %>% 
  count(remus) 


print(paste("Ratio of HH receiving remittances if one member is abroad",
            remittances_reciving_ratio$n[1]/(remittances_reciving_ratio$n[1]+ 
                                               remittances_reciving_ratio$n[2])))

```






# Using remittances and purpose of remittances

```{r}
remit_purposes <- mig_filtered %>% 
  mutate(remit1 = remit2) %>% 
  filter(remit1 != 8888 & remit != 9999) %>% 
  mutate(remit_1_mod = case_when(remit1 == 1 ~ "consmp",
                               remit1 == 2 ~ "consmp",
                               remit1 == 7 ~ "invest",
                               remit1 == 6 ~ "invest"
                               
                        )) %>% 
    filter(remit_1_mod == "consmp" | remit_1_mod == "invest") 



house_filtered_remit_usage <- house_filtered %>% 
  left_join(remit_purposes, by = c("commun" = "commun", "hhnum" = "hhnum", "surveyyr" = "surveyyr")) %>% 
  drop_na(remit_1_mod) %>% 
  mutate(remit_2_mod = ifelse(remit_1_mod == "consmp", 0, 1))

m1_remittances_usage <- glm(remit_2_mod  ~  total_land, data = house_filtered_remit_usage, family = "binomial")

stargazer(m1_remittances_usage , type = "text")



```





Calculate the share of correct predictions
```{r}
predicted_probs <- expand.grid(total_land= seq(0,10))
estimates <- augment(m1_remittances_usage, type.predict = "response", newdata = predicted_probs)
ggplot(data = estimates, aes(x = total_land, y = .fitted))+
  geom_line()+
  labs(title="Usage of Remittances as a function of land size")+
  theme_bw()+
  ylab("Predicted probability")+
  xlab("Land [ha]")
```












# Descriptives


### Land Distribution

```{r}



calculate_land <- house_filtered %>% 
  mutate(clhect2 = ifelse(clhect2 == 8888 | clhect2 == 9999, 0, clhect2)) %>% 
  mutate(clhect3 = ifelse(clhect3 == 8888 | clhect3 == 9999, 0, clhect3)) %>% 
  mutate(clhect4 = ifelse(clhect4 == 8888 | clhect4 == 9999, 0, clhect4)) %>% 
  mutate(total_land = clhect1 + clhect2 + clhect3, clhect4)

ggplot(data = calculate_land, aes(x = total_land))+
  geom_histogram(binwidth = 1)+
  xlim(0,15)+
  labs(title = "Distribution of Land [ha] within sampled households ")+
  xlab("Land in hectars")+
  ylab("Number of households")+
  theme_bw()+
  geom_vline(aes(xintercept = median(calculate_land$total_land)), color = "red", linetype = "dashed")

```




### Children
```{r}
life_filtered %>% 
  filter(age > 45 & cunion == 1) %>% 
  ggplot(aes(x = ceb))+
  geom_histogram()+
  geom_vline(aes(xintercept = median(ceb)), color = "red")


```












# Script to create sample dataset for ABM on individuals and households


```{r}
commun <- read.csv("commun174.csv")
house <- read.csv("house170.csv")
pers <- read.csv("pers170.csv")

commun_filtered <- commun %>% 
  filter(COMPOP90 < 2500) %>% #less than 5000 inhabitants in community --> rural
  filter(AGRIM90 > .5) # more than 50% of men work in agriculture --> agrarian

rural_agrarian_communities_states <- read_delim("rural_agrarian_communities_states.csv", 
                                                ";", escape_double = FALSE, trim_ws = TRUE) 

rural_agrarian_communities_states <- rural_agrarian_communities_states %>% 
  dplyr::filter(state %in% c("Guerrero", "Oaxaca", "Jalisco", "Michoacán de Ocampo"))
                                     
pers_filtered <- pers %>% 
  semi_join(house_filtered, by = c("commun" = "commun", "hhnum" = "hhnum", "surveyyr" = "surveyyr"))

house_filtered <- house %>% 
  dplyr::semi_join(rural_agrarian_communities_states, by = c("commun" = "commun")) %>%  #investigate only HH in rural agrarian communities
  dplyr::filter(famw != 8888 & famw != 9999 & famw != 0) %>% 
  dplyr::filter(clhect1 != 8888 & clhect1 != 9999) %>% #use only those who own land
  dplyr::filter(clhect2 == 8888 | clhect2 == 9999 | clhect2 < 10) %>% 
  dplyr::filter(clhect3 == 8888 | clhect3 == 9999 | clhect3 < 10) %>% 
  dplyr::filter(clhect4 == 8888 | clhect4 == 9999 | clhect4 < 10) %>% 
  dplyr::filter(clhect1 < 10) %>%    #aim for smallholder agriculture
  mutate(clhect2 = ifelse(clhect2 == 8888 | clhect2 == 9999, 0, clhect2)) %>% 
  mutate(clhect3 = ifelse(clhect3 == 8888 | clhect3 == 9999, 0, clhect3)) %>% 
  mutate(clhect4 = ifelse(clhect4 == 8888 | clhect4 == 9999, 0, clhect4)) %>% 
  mutate(total_land = clhect1 + clhect2 + clhect3, clhect4)


#Draw Sample of Land Distribution
household_sample <- house_filtered %>% 
  dplyr::filter(total_land <= 10) %>% 
  sample_n(225) %>%  # to align 15*15 agents on grid 
  select(total_land, lifecycl, members, workers, nonwork, commun, hhnum, surveyyr)


person_household_sample <- pers_filtered %>% 
  semi_join(household_sample, by = c("commun" = "commun", "hhnum" = "hhnum", "surveyyr" = "surveyyr")) %>% 
  mutate(migrated_us = ifelse(usyr1 == 8888, 0, 1)) %>% 
  dplyr::filter(age != 8888) %>% #small number of cases, though filter for outliers here
  select(relhead, commun, hhnum, surveyyr, age, sex) %>% 
  dplyr::filter(relhead  <= 3) #for simplification we yet only use families of head, spouse, children


#write_csv(household_sample, "household_sample.csv")
#write_csv(person_household_sample, "household_members_sample.csv")
```


