---
title: "MMP Data Analysis and Models"
author: "Jan Sodoge"
date: "3/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stargazer)
library(broom)
library(DescTools)
library(ggsci)
library(ggsci)
```


# Create Sample
Following the criteria outlined in the thesis the dataset is created based on the full 
MMP dataset. 
```{r,  echo=FALSE}
#comun data
commun <- read.csv("mmp_data/commun174.csv")
house <- read.csv("mmp_data/house170.csv")
life <- read.csv("mmp_data/life170.csv")

pers <- read.csv("mmp_data/pers170.csv")
mig <- read.csv("mmp_data/mig170.csv")


commun_filtered <- commun %>% 
  filter(COMPOP90 < 2500) %>% #less than 5000 inhabitants in community --> rural
  filter(AGRIM90 > .5) # more than 50% of men work in agriculture --> agrarian


house_filtered <- house %>% 
  dplyr::semi_join(commun_filtered, by = c("commun" = "COMMUN")) %>%  #investigate only HH in rural agrarian communities
  dplyr::filter(famw != 8888 & famw != 9999 & famw != 0) %>% 
  dplyr::filter(clhect1 != 8888 & clhect1 != 9999) %>% #use only those who own land
  dplyr::filter(clhect2 == 8888 | clhect2 == 9999 | clhect2 < 10) %>% 
  dplyr::filter(clhect3 == 8888 | clhect3 == 9999 | clhect3 < 10) %>% 
  dplyr::filter(clhect4 == 8888 | clhect4 == 9999 | clhect4 < 10) %>% 
  dplyr::filter(clhect1 < 10) %>%   #aim for smallholder agriculture
  mutate(clhect2 = ifelse(clhect2 == 8888 | clhect2 == 9999, 0, clhect2)) %>% 
  mutate(clhect3 = ifelse(clhect3 == 8888 | clhect3 == 9999, 0, clhect3)) %>% 
  mutate(clhect4 = ifelse(clhect4 == 8888 | clhect4 == 9999, 0, clhect4)) %>% 
  mutate(total_land = clhect1 + clhect2 + clhect3, clhect4)




life_filtered <- life %>% 
  semi_join(house_filtered,  by = c("commun" = "commun", "hhnum" = "hhnum", "surveyyr" = "surveyyr")) 
  
pers_filtered <- pers %>% 
  semi_join(house_filtered, by = c("commun" = "commun", "hhnum" = "hhnum", "surveyyr" = "surveyyr"))


mig_filtered <- mig %>% 
  semi_join(house_filtered, by = c("commun" = "commun", "hhnum" = "hhnum", "surveyyr" = "surveyyr"))

```







# Descriptive Statistics

```{r}
#household structure
p_1 <- house_filtered %>% 
  select(members, workers, nonwork,
    sons, daus, sonslabf, sodomig, sousmig,
         daus, dauslabf, dusmig, ddomig,
         remus) %>% 
  mutate(sonslabf = ifelse(sonslabf == 8888, NA, sonslabf),
         sodomig = ifelse(sodomig == 8888, NA, sodomig),
         sousmig = ifelse(sousmig == 8888, NA, sousmig),
         dauslabf = ifelse(dauslabf == 8888, NA, dauslabf),
         dusmig = ifelse(dusmig == 8888, NA, dusmig),
         ddomig = ifelse(ddomig == 8888, NA, ddomig),
         remus = ifelse(remus == 1,1,0)
  ) %>% 
  stargazer::stargazer(summary = TRUE, type="text",
                        summary.stat = c("N", "median", "mean", "sd"))



person_household_sample <- pers_filtered %>% 
  mutate(migrated_us = ifelse(usyr1 == 8888, 0, 1),
         currently_migrated = ifelse(uscurtrp == 1 | uscurtrp == 2, 1,0)) %>% 
  dplyr::filter(age != 8888) %>% #small number of cases, though filter for outliers here
  select(relhead, commun, hhnum, surveyyr, age, sex, currently_migrated) %>% 
  dplyr::filter(relhead  <= 3) %>% 
  group_by(hhnum, surveyyr, commun) %>% 
  count(relhead) %>% 
  ungroup() %>% 
  select(relhead, n) %>% 
 pivot_wider(names_from = relhead, values_from = n) 



migrants_per_hh <- pers_filtered %>% 
  mutate(currently_migrated = ifelse(uscurtrp == 1, 1,0)) %>% 
  group_by(commun, surveyyr, hhnum) %>% 
  summarise(us_migrants_per_household = sum(currently_migrated)) %>% 
  group_by(surveyyr) %>% 
  summarise(mean_migrants_per_hh  = mean(us_migrants_per_household, na.rm = TRUE)) 

```



```{r}

#Land Distribution Plot
calculate_land <- house_filtered %>% 
  mutate(clhect2 = ifelse(clhect2 == 8888 | clhect2 == 9999, 0, clhect2)) %>% 
  mutate(clhect3 = ifelse(clhect3 == 8888 | clhect3 == 9999, 0, clhect3)) %>% 
  mutate(clhect4 = ifelse(clhect4 == 8888 | clhect4 == 9999, 0, clhect4)) %>% 
  mutate(total_land = clhect1 + clhect2 + clhect3, clhect4)

 ggplot(data = calculate_land, aes(x = total_land))+
  geom_histogram(binwidth = 1)+
     annotate(geom = "text", x = 3, y  = 300, label = paste("Median:", median(calculate_land$total_land)), color = "red")+

  xlim(0,15)+
  xlab("Land in hectars")+
  ylab("Number of households")+
  theme_minimal()+
  geom_vline(aes(xintercept = median(total_land)), color = "red", linetype = "dashed")

```




```{r}
p3 <- pers_filtered %>% 
  mutate(ustrips = ifelse(ustrips ==  0, 0,1)) %>% 
  group_by(surveyyr) %>% 
 summarise(migrants = mean(ustrips, na.rm = TRUE)) %>% 
  ggplot(aes(x = surveyyr, y = migrants))+
  geom_point()+
  geom_smooth(method=lm)+
  #xlim(c(1995,2015))+
  labs(subtitle = "pers-file based")
```








# Models predciting migration US
```{r}



life_filtered_us_migrants <- life_filtered %>% 
  dplyr::filter(usyr1 != 8888) %>% 
  filter(usyr1 == year) %>% 
  select(usyr1, fausmig, mousmig, brousmig, sisusmig, land, hectars, age, commun, children) %>% 
  filter(age > 18)


life_filtered_non_migrants <- life_filtered %>% 
  dplyr::filter(usyr1 == 8888) %>% 
  select(usyr1, fausmig, mousmig, brousmig, sisusmig, land, hectars, age, commun, children) %>% 
  filter(age > 18)


age_data_migrants <- as.vector(life_filtered_us_migrants$age)

counterfactual_ages <- data.frame()
for(migrant_age in age_data_migrants){
  tmp <- dplyr::filter(life_filtered_non_migrants, age == migrant_age) 
  
  
  counterfactual_ages <- dplyr::bind_rows(counterfactual_ages, tmp)
  
  
}

counterfactual_ages <- counterfactual_ages %>% 
  distinct()



dataset <- dplyr::bind_rows(life_filtered_us_migrants, counterfactual_ages) %>% 
  mutate(migrated_us = ifelse(usyr1 == 8888, 0, 1),
         social_network = fausmig + mousmig, brousmig + sisusmig,
         children = ifelse(children > 0, 1,0 )) 
  

m1_a <- glm(migrated_us ~ social_network + land, data = dataset, family = "binomial")
m1_b <- glm(migrated_us ~ social_network + hectars, data = dataset, family = "binomial")

m2_a <- glm(migrated_us ~ social_network + land + log(age) + children , data = dataset, family = "binomial")
m2_b <- glm(migrated_us ~ social_network + hectars + log(age) + children , data = dataset, family = "binomial")

m3_a <- glm(migrated_us ~ social_network + land + log(age)+ children +  as.factor(commun), data = dataset, family = "binomial")
m3_b <- glm(migrated_us ~ social_network + hectars + log(age)+ children +  as.factor(commun), data = dataset, family = "binomial")


library(bife)
m3_a_bife <-  bife(migrated_us ~ social_network + land + children + log(age)  | commun, dataset, "probit")
m3_b_bife <-  summary(bife(migrated_us ~ social_network + hectars + children + log(age)  | commun, dataset, "probit"))

stargazer(m1_a, m1_b, m2_a, m2_b, m3_a, m3_b, type = "text")


```


```{r}
cat(paste("Derived coefficients used for NetLogo agent-behavior paramtrizstion",
            "
          For intercept", coef(m2_b)[['(Intercept)']],
            "
          For Children", coef(m2_b)[["children"]],
            "
          For social network", coef(m2_b)[["social_network"]],
            "
          For land", coef(m2_b)[["hectars"]],
            "
          For log(age)", coef(m2_b)[["log(age)"]]))
```


Within the Netlogo model this translates to the following structural form to derive a migration probability (which needs to be re-adjusted for out-migration statistics). From the logistic regression above, we derive the following coefficients: 

$\beta_1$ age;
$\beta_2$ children;
$\beta_3$ hectares;
$\beta_4$ network-effect;
$\beta_5$ intercept; 


$$p = exp(\beta_5 + (\beta_1 * log(age)) + (\beta_3 * land) + (\beta_4 * network)) / (1+ exp(\beta_5 + (\beta_1 * log(age)) + (\beta_3 * land) + (\beta_4 * network)))$$



Derive a prediction formula for the above-specified parameters
```{r}
predicted_probs <- expand.grid(age= 20, hectars= seq(0,10), social_network = seq(0,1), children = 0)
estimates <- augment(m2_b, type.predict = "response", newdata = predicted_probs)
estimates <- estimates %>% 
  mutate(social_network = as.factor(social_network))
ggplot(data = estimates, aes(x = hectars, y = .fitted, fill = social_network, color = social_network))+
  geom_line()+
   theme_minimal()+
  scale_color_uchicago()+
  theme(legend.position = "bottom")+
  ylab("Predicted Probability")+
  xlab("Land [hectares]")+
  labs(colour="Migration Experience in Social Network")
```
```{r}

prediction_data <- augment(m2_b, type.predict = "response") %>% 
  mutate(.pred = as.numeric(.fitted > .5))
mean(prediction_data$migrated_us == prediction_data$.pred, na.rm = TRUE)
```






# Models predciting migration Domestic
```{r}
life_filtered_us_migrants <- life_filtered %>% 
  dplyr::filter(doyr1 != 8888) %>% 
  filter(doyr1 == year) %>% 
  select(doyr1, fausmig, mousmig, brousmig, sisusmig, land, hectars, age, commun, children) %>% 
  filter(age > 18)


life_filtered_non_migrants <- life_filtered %>% 
  dplyr::filter(doyr1 == 8888 & usyr1 == 8888) %>% 
  select(doyr1, fausmig, mousmig, brousmig, sisusmig, land, hectars, age, commun, children) %>% 
  filter(age > 18)


age_data_migrants <- as.vector(life_filtered_us_migrants$age)

counterfactual_ages <- data.frame()
for(migrant_age in age_data_migrants){
  tmp <- dplyr::filter(life_filtered_non_migrants, age == migrant_age) 
  counterfactual_ages <- dplyr::bind_rows(counterfactual_ages, tmp)
}

counterfactual_ages <- counterfactual_ages %>% 
  distinct()

dataset <- dplyr::bind_rows(life_filtered_us_migrants, counterfactual_ages) %>% 
  mutate(migrated_do = ifelse(doyr1 == 8888, 0, 1),
         social_network = fausmig + mousmig, brousmig + sisusmig, # we lack these variables for the US
         children = ifelse(children > 0, 1,0 )) 
  

m1_a_do <- glm(migrated_do ~   land, data = dataset, family = "binomial")
m1_b_do <- glm(migrated_do ~  hectars, data = dataset, family = "binomial")
m2_a_do <- glm(migrated_do ~  land + log(age) + children , data = dataset, family = "binomial")
m2_b_do <- glm(migrated_do ~  hectars + log(age) + children , data = dataset, family = "binomial")
m3_a_do <- glm(migrated_do ~  land + log(age)+ children +  as.factor(commun), data = dataset, family = "binomial")
m3_b_do <- glm(migrated_do ~  hectars + log(age)+ children +  as.factor(commun), data = dataset, family = "binomial")


stargazer(m1_a_do, m1_b_do, m2_a_do, m2_b_do, m3_a_do, m3_b_do, type = "text")


```










# Out-migration rates


What is the share of individuals over time who have migration experience?


```{r}
pers_filtered %>% 
  mutate(ustrips = ifelse(ustrips == 0 & dotrips == 0, 0,1)) %>% 
  group_by(surveyyr) %>% 
 summarise(migrants = mean(ustrips)) %>% 
  ggplot(aes(x = surveyyr, y = migrants))+
  geom_point()+
  geom_smooth(method=lm)+
  xlim(c(1995,2015))
```





# Using remittances and purpose of remittances

```{r}
remit_purposes <- mig_filtered %>% 
  mutate(remit1 = remit2) %>% 
  filter(remit1 != 8888 & remit != 9999) %>% 
  mutate(remit_1_mod = case_when(remit1 == 1 ~ "consmp",
                               remit1 == 2 ~ "consmp",
                               remit1 == 7 ~ "invest",
                               remit1 == 6 ~ "invest"
                               
                        )) %>% 
    filter(remit_1_mod == "consmp" | remit_1_mod == "invest") 



house_filtered_remit_usage <- house_filtered %>% 
  left_join(remit_purposes, by = c("commun" = "commun", "hhnum" = "hhnum", "surveyyr" = "surveyyr")) %>% 
  drop_na(remit_1_mod) %>% 
  mutate(remit_2_mod = ifelse(remit_1_mod == "consmp", 0, 1))

m1_remittances_usage <- glm(remit_2_mod  ~  total_land, data = house_filtered_remit_usage, family = "binomial")

stargazer(m1_remittances_usage , type = "text")



```





```{r}



cat(paste("Derived coefficients used for NetLogo agent-behavior paramtrizstion",
            "
          For intercept", coef(m1_remittances_usage)[['(Intercept)']],
            "
          For Children", coef(m1_remittances_usage)[["total_land"]]))



```




Calculate the share of correct predictions
```{r}
predicted_probs <- expand.grid(total_land= seq(0,10))
estimates <- augment(m1_remittances_usage, type.predict = "response", newdata = predicted_probs)
ggplot(data = estimates, aes(x = total_land, y = .fitted))+
  geom_line()+
  labs(title="Usage of Remittances as a function of land size")+
  theme_bw()+
  ylab("Predicted probability")+
  xlab("Land [ha]")
```
















### Children
We use this distribution to calibrate the birth probability per spouse per year 
```{r}
life_filtered %>% 
  filter(age > 45 & cunion == 1) %>% 
  ggplot(aes(x = ceb))+
  geom_histogram()+
  geom_vline(aes(xintercept = median(ceb)), color = "red")


```















# Script to create sample dataset for ABM on individuals and households


```{r}


commun <- read.csv("mmp_data/commun174.csv")
house <- read.csv("mmp_data/house170.csv")
pers <- read.csv("mmp_data/pers170.csv")

commun_filtered <- commun %>% 
  filter(COMPOP90 < 2500) %>% #less than 5000 inhabitants in community --> rural
  filter(AGRIM90 > .5) # more than 50% of men work in agriculture --> agrarian

rural_agrarian_communities_states <- read_delim("rural_agrarian_communities_states.csv", 
                                                ";", escape_double = FALSE, trim_ws = TRUE) 


rural_agrarian_communities_states <- rural_agrarian_communities_states %>% 
  dplyr::filter(state %in% c("Guerrero", "Oaxaca", "Jalisco", "Michoacán de Ocampo"))
                                     
                                  
pers_filtered <- pers %>% 
  semi_join(house_filtered, by = c("commun" = "commun", "hhnum" = "hhnum", "surveyyr" = "surveyyr"))



house_filtered <- house %>% 
  dplyr::semi_join(rural_agrarian_communities_states, by = c("commun" = "commun")) %>%  #investigate only HH in rural agrarian communities
  dplyr::filter(famw != 8888 & famw != 9999 & famw != 0) %>% 
  dplyr::filter(clhect1 != 8888 & clhect1 != 9999) %>% #use only those who own land
  dplyr::filter(clhect2 == 8888 | clhect2 == 9999 | clhect2 < 10) %>% 
  dplyr::filter(clhect3 == 8888 | clhect3 == 9999 | clhect3 < 10) %>% 
  dplyr::filter(clhect4 == 8888 | clhect4 == 9999 | clhect4 < 10) %>% 
  dplyr::filter(clhect1 < 10) %>%    #aim for smallholder agriculture
  mutate(clhect2 = ifelse(clhect2 == 8888 | clhect2 == 9999, 0, clhect2)) %>% 
  mutate(clhect3 = ifelse(clhect3 == 8888 | clhect3 == 9999, 0, clhect3)) %>% 
  mutate(clhect4 = ifelse(clhect4 == 8888 | clhect4 == 9999, 0, clhect4)) %>% 
  mutate(total_land = clhect1 + clhect2 + clhect3, clhect4)


#Draw Sample of Land Distribution
household_sample <- house_filtered %>% 
  dplyr::filter(total_land <= 10) %>% 
  sample_n(225) %>%  # to align 15*15 agents on grid 
  select(total_land, lifecycl, members, workers, nonwork, commun, hhnum, surveyyr)


person_household_sample <- pers_filtered %>% 
  semi_join(household_sample, by = c("commun" = "commun", "hhnum" = "hhnum", "surveyyr" = "surveyyr")) %>% 
  mutate(migrated_us = ifelse(usyr1 == 8888, 0, 1)) %>% 
  dplyr::filter(age != 8888) %>% #small number of cases, though filter for outliers here
  select(relhead, commun, hhnum, surveyyr, age, sex) %>% 
  dplyr::filter(relhead  <= 3) #for simplification we yet only use families of head, spouse, children


#write_csv(household_sample, "household_sample.csv")
#write_csv(person_household_sample, "household_members_sample.csv")





  
  

```




```{r}
households <- read.csv("household_sample.csv")
members <- read.csv("household_members_sample.csv")
```


# Generate overview of existing inequality levels´


Utility is defined as the agricultural land
```{r}
ggplot(data = households, aes(x = total_land))+
  geom_histogram(binwidth = 1, color = "white")+
  annotate(geom="text",  x=7.5, y=50, label= paste("Gini Index: ", round(DescTools::Gini(households$total_land),2)))+
  theme_classic()


```





Relative utility is defined as agricultural land / number of persons per house
```{r}


hh_member <- members %>% 
  group_by(commun, hhnum, surveyyr) %>% 
  count() 

relative_util <- households %>% 
  left_join(hh_member) %>% 
  mutate(relative_utility = total_land / n) 
households %>% 
  left_join(hh_member) %>% 
  mutate(relative_utility = total_land / n) %>% 
  ggplot(aes(x= relative_utility))+
  geom_histogram()+
  annotate(geom="text",  x=3, y=50, label= paste("Gini Index: ", round(DescTools::Gini(relative_util$relative_utility, na.rm = TRUE),2)))+
  theme_classic()
  

```





